\texttt{davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2022-11-16 paper/main.tex
finished illustrative example section

Next, the user smuggles a pair of packages that extend the Jupyter notebook interface with various interactive JavaScript widgets. The \texttt{ipywidgets}~\cite{GrouEtal15} package provides an API for creating these widgets with Python code, and the \texttt{widgetsnbextension} package provides the machinery for the notebook frontend to display them. Thus, \texttt{ipywidgets} must be installed in the same environment as the IPython kernel, and \texttt{widgetsnbextension} must be installed in the environment that houses the Jupyter notebook server. In the simplest case, these two environments are the same; however, a common approach to using Jupyter notebooks with multiple different Python environments (e.g., for different tasks or projects) is to run the notebook server from a ``base environment,'' with additional environments each providing their own separate, interchangeable IPython kernels. (While possible to detect programmatically, Figure~\ref{fig:illustrative-example} assumes this setup to simplify the example code.) To handle this multi-environment installation, line 25 temporarily changes the \texttt{pip} executable used to install smuggled packages to that of the notebook server's environment before smuggling \texttt{widgetsnbextension} (line 26). Since there is no need to actually load this package for later use in the analysis code, it is aliased as a single underscore---by convention, a ``throwaway'' variable in Python. In line 27, the user changes the \texttt{pip} executable back to that of the notebook kernel environment (its original value) before installing \texttt{ipywidgets} and subsequent packages. With these two packages installed, line 30 smuggles \texttt{tqdm}~\cite{CaspEtal22}, which provides convenient progress bars for long-running code. In Jupyter notebooks, the \texttt{tqdm.notebook} module can be imported to enable prettier progress bars displayed via \texttt{ipywidgets}, if that package is installed and importable. Therefore, in this case, it was important to \texttt{smuggle}

% In Jupyter notebooks, \texttt{tqdm} will attempt to display these progress bars using \texttt{ipywidgets} if the package is installed and available.

%managing multiple, interchangeable IPython kernels is to run the notebook server from a ``base environment,'' with additional Python environments each providing its own separate kernel (e.g.,

Finally, the user loads in the ... but in order to successfully ... they must deal with a similar problem to ...

>>>>

Next, the user smuggles a pair of packages that extend the Jupyter notebook interface with various interactive JavaScript widgets. The \texttt{ipywidgets}~\cite{GrouEtal15} package provides an API for creating these widgets with Python code, and the \texttt{widgetsnbextension} package provides the machinery for the notebook frontend to display them. Thus, \texttt{ipywidgets} must be installed in the same environment as the IPython kernel, and \texttt{widgetsnbextension} must be installed in the environment that houses the Jupyter notebook server. In the simplest case, these two environments are the same; however, a common approach to using Jupyter notebooks with multiple different Python environments (e.g., for different tasks or projects) is to run the notebook server from a ``base environment,'' with additional environments each providing their own separate, interchangeable IPython kernels. (While possible to detect programmatically, Figure~\ref{fig:illustrative-example} assumes this setup to simplify the example code.) To handle this multi-environment installation, line 25 temporarily sets the \texttt{pip} executable used to install smuggled packages to that of the notebook server's environment before smuggling \texttt{widgetsnbextension} (line 26). Since there is no need to actually load this package for later use in the analysis code, it is aliased as a single underscore---by convention, a ``throwaway'' variable in Python. In line 27, the user changes the \texttt{pip} executable back to that of the notebook kernel environment (its original value) before installing \texttt{ipywidgets} and subsequent packages. With these two packages installed, line 30 smuggles \texttt{tqdm}~\cite{CaspEtal22}, which provides convenient progress bars for long-running code. In Jupyter notebooks, the \texttt{tqdm.notebook} module can be imported to enable prettier progress bars displayed via \texttt{ipywidgets}, if that package is installed and importable. Therefore, to take advantage of this feature, it was important to \texttt{smuggle} \texttt{tqdm} after ensuring the \texttt{ipywidgets} package was available.

Finally, the user loads in the re-formatted dataset (line 32) and pre-trained model (line 34) they'll use in their analyses. But in accessing the latter, the user encounters an issue similar to the one they initially faced with the dataset. This hypothetical semantic model was provided as a \texttt{scikit-learn} ``\texttt{Pipeline}'' object that allows the user to pass text data through two pre-trained transformers, in series: a \texttt{CountVectorizer} instance and a \texttt{LatentDirichletAllocation} instance (i.e., a topic model \cite{BleiEtal03}). The model was saved by its original creator using the \texttt{joblib} library, as \texttt{scikit-learn}'s documentation recommends, with the caveat that because \texttt{joblib} uses the \texttt{pickle} protocol internally, the ability to save and load pre-trained models is not guaranteed across versions due to \texttt{pickle}'s requirement that referenced classes be importable from the same location when saved and loaded. This issue affects the user's model by way of the \texttt{LatentDirichletAllocation} class---when the \texttt{Pipeline} object was created (again, in June, 2019, using \texttt{scikit-learn} v0.21.3), the class's definition lived in the \texttt{sklearn.decomposition.online\_lda} module. However, in version 0.22.0 (released November, 2019), that module was renamed to ``\texttt{\_online\_lda},'' and in v0.22.1 (January, 2020) it was again renamed to ``\texttt{\_lda}'' (which has remained its name since then). Thus, in order to successfully load the model that includes the pre-trained \texttt{LatentDirichletAllocation} instance, in line 33, the user first smuggles a version of \texttt{scikit-learn} prior to v0.22.0 (i.e., before the first time the module's name was changed). Once the model is loaded and reconstructed in memory  from a compatible package version (line 34), the user then upgrades to a newer version of \texttt{scikit-learn} for use elsewhere in their analyses. When 
line 35 is run, \texttt{davos} will replace the existing \texttt{scikit-learn} version with v1.1.3 in both the user's local package environment and in memory without interruption, such that ``\texttt{sklearn}'' will reference the new version in any code below. However, this will not affect the already compiled code object stored in the ``\texttt{transformer}'' variable, meaning that the pre-trained model will be usable simultaneously. 
% It's possible that in some cases, a version mismatch like this could introduce certain compatibility issues, depending on the specific package and the differences between the two versions being used. For example, a function from a newer package version could fail to handle an object from an older version passed to it as an argument, if it expects the object to have certain attributes that had not yet been added in the older version. Fortunately, with \texttt{davos}, this scenario is relatively easy to handle, as \texttt{smuggle} statements can be used to temporarily switch between package versions as needed (e.g., for a single line, in order to call an older version of the function). And because the installer program caches previously installed package versions, switching between package versions in this way can be done fairly quickly.
